aircraft-pfp

### 2025-04-01 README.Rmd
readme
add capability to read CO2 and potentially other gases

The first step consists in constructing a summary for the ObsPack.
This is required to read the data, but also, identify `agl`, 
which is present in some of the file names.
This function returns a `data.frame`.
Optionally, the user can indicate a path to store the `data.frame`.
`obs_summary` also prints a summary of the data. The second
argument is the categories, and by default includes the categories
shown below, to account for all the files.
Then the summary `data.frame` contains 
the columns `id` as the full path to each
file, `name` which is the name or relative path of the file, `n`
just an id, `sector` such as tower, and the column `agl`
which indicates the `agl` indicated in the name of the file
if available. To read the documentation of this function,
the user must run `?obs_summary`.

@@

@@

## Special thanks to all the contributors

@@

>>>>

The first step consists in constructing a summary for ObsPack (CH4, CO2 or other).
This is required to read the data, but also, identify `agl`, 
which is present in some of the file names.
This function returns a `data.frame`.
Optionally, the user can indicate a path to store the `data.frame`.
`obs_summary` also prints a summary of the data. The second
argument is the categories, and by default includes the categories
shown below, to account for all the files.
Then the summary `data.frame` contains 
the columns `id` as the full path to each
file, `name` which is the name or relative path of the file, `n`
just an id, `sector` such as tower, and the column `agl`
which indicates the `agl` indicated in the name of the file
if available. To read the documentation of this function,
the user must run `?obs_summary`.

@@

e.g. for the `aircraft-pfp` we can read $CO_2$ and $CH_4$ [here](https://noaa-gml.github.io/rtorf/articles/aircraft-pfp.html)

@@

## Special thanks to all the

**contributors**

@@

and

**Stargazers**

<p>
  <a href="https://github.com/noaa-gml/rtorf/stargazers">
    <img src="http://reporoster.com/stars/dark/noaa-gml/rtorf"/>
  </a>
</p>

### 2025-04-01 examples/aircraft-pfp.Rmd
add capability to read CO2 and potentially other gases

obs <- "Z:/obspack/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/"
index <- obs_summary(obs = obs, 
                     categories = cate)
```

@@

now we can cut time every 20 seconds. We can chosse other frequency as well.

@@

```{r addlt, eval = T}
df3 <- obs_addltime(df2)
setorderv(df3, cols = c("site_code", "timeUTC"),
          order = c(-1, 1))
df3

@@

```{r csvy, eval = T}
cat("\nAdding notes in csvy:\n")
notes <- c(paste0("sector: ", datasetid),
           paste0("timespan: ", yy),
           paste0("spatial_limits: north = ", north, ", south = ", south, ", east = ", east, ", west = ", west),
           "data: Data averaged every 20 seconds",
           paste0("altitude: < ", max_altitude),
           "hours: All",
           "local_time: if var `site_utc2lst` is not available, calculated as",
           "longitude/15*60*60 (John Miller)")

@@

```{r obs_plotsave, fig.width=5, fig.height=3, eval = T, echo = F, message=F, warning=F}
png("../man/figures/obsplot_aircraftpfp.png", width = 1500, height = 1000, res = 200)
obs_plot(df3, time = "timeUTC", yfactor = 1e9)
dev.off()
```

```{r obs_plot, fig.width=5, fig.height=3, eval = F}
obs_plot(df3, time = "timeUTC", yfactor = 1e9)
```

@@

```{r savesf, fig.width=5, fig.height=3, eval = T, echo = F}
library(sf)
x <- st_as_sf(df3, coords = c("longitude", "latitude"), crs = 4326)
png("../man/figures/obsplot_aircraftpfp_map.png", width = 1500, height = 1000, res = 200)
plot(x["value"], axes = T, reset = F)
maps::map(add = T)
dev.off()
```

```{r sf, fig.width=5, fig.height=3, eval = F}
library(sf)
x <- st_as_sf(df3, coords = c("longitude", "latitude"), crs = 4326)
plot(x["value"], axes = T, reset = F)
maps::map(add = T)
```

@@

```{r vertplot, fig.width=5, fig.height=3, eval = T, echo = F, message=F, warning=F}
x <- df3
x$ch4 <- x$value*1e+9
png("../man/figures/obsplot_aircraftpfp_vert.png", width = 1500, height = 1000, res = 200)
obs_plot(x, 
         time = "ch4", 
         y = "altitude_final", 
         colu = "month", n = c(1L, 3L, 6L, 8L, 9L, 11L, 12L), 
         type = "b", 
         xlab = expression(CH[4]~ppb), 
         ylab = "altitude (m)")
dev.off()
```

```{r obs_plot_vertt, fig.width=7, fig.height=5, eval = F}
x <- df3
x$ch4 <- x$value*1e+9
obs_plot(x, 
         time = "ch4", 
         y = "altitude_final", 
         colu = "site_code", #n = c(1L, 3L, 6L, 8L, 9L, 11L, 12L), 
         type = "b", 
         xlab = expression(CH[4]~ppb), 
         ylab = "altitude (m)")
```

>>>>

obsch4 <- "Z:/obspack/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/"
indexch4 <- obs_summary(obs = obsch4, 
                        categories = cate)
```

How let us read CO2

```{r}
obsco2 <- "Z:/obspack/obspack_co2_1_GLOBALVIEWplus_v9.1_2023-12-08/data/nc/"
indexco2 <- obs_summary(obs = obsco2, 
                        categories = cate)
```

We can merge co2 and ch4

```{r}
index <- rbind(indexch4,
               indexco2)
```

@@

now we can cut time every 20 seconds. We can choose other frequency as well.

@@

```{r addlt, eval = T}
df3 <- obs_addltime(df2)
setorderv(df3, cols = c("dataset_calibration_scale", "site_code", "timeUTC"),
          order = c(1, -1, 1))
df3

@@

```{r csvy, eval = T}
cat("\nAdding notes in csvy:\n")
notes <- c(paste0("sector: ", datasetid),
           paste0("timespan: ", yy),
           paste0("spatial_limits: north = ", north, ", south = ", south, ", east = ", east, ", west = ", west),
           "gas: CO2 and CH4",
           "data: Data averaged every 20 seconds",
           paste0("altitude: < ", max_altitude))

@@

```{r obs_plotsave, fig.width=5, fig.height=3, eval = T, echo = F, message=F, warning=F}
png("../man/figures/obsplot_aircraftpfp.png", 
    width = 1500,
    height = 1000,
    res = 200)
obs_plot(df3[dataset_calibration_scale == "WMO CO2 X2019"],
         time = "timeUTC",
         yfactor = 1e6)
dev.off()
```

```{r obs_plot, fig.width=5, fig.height=3, eval = F}
obs_plot(df3[dataset_calibration_scale == "WMO CO2 X2019"],
         time = "timeUTC",
         yfactor = 1e6)
```

@@

```{r savesf, fig.width=5, fig.height=3, eval = T, echo = F}
library(sf)
x <- df3[dataset_calibration_scale == "WMO CO2 X2019",
         mean(value*1e+6),
         by = .(site_code,
                latitude,
                longitude)]
x <- st_as_sf(x, 
              coords = c("longitude",
                         "latitude"), 
              crs = 4326)
png("../man/figures/obsplot_aircraftpfp_map.png",
    width = 1500,
    height = 1000,
    res = 200)
plot(x["V1"],
     axes = T,
     reset = F)
maps::map(add = T)
dev.off()
```

```{r sf, fig.width=5, fig.height=3, eval = F}
library(sf)
x <- df3[dataset_calibration_scale == "WMO CO2 X2019",
         mean(value*1e+6),
         by = .(site_code,
                latitude,
                longitude)]
x <- st_as_sf(x, 
              coords = c("longitude",
                         "latitude"), 
              crs = 4326)
plot(x["V1"], 
     axes = T, 
     reset = F)
maps::map(add = T)
```

@@

```{r vertplot, fig.width=5, fig.height=3, eval = T, echo = F, message=F, warning=F}
x <- df3[dataset_calibration_scale == "WMO CO2 X2019"]
x$co2 <- x$value*1e+6
png("../man/figures/obsplot_aircraftpfp_vert.png", 
    width = 1500,
    height = 1000,
    res = 200)
obs_plot(x, 
         time = "co2",
         y = "altitude_final", 
         colu = "month", 
         type = "p", 
         xlab = expression(CO[2]~ppm), 
         ylab = "altitude (m)")
dev.off()
```

```{r obs_plot_vertt, fig.width=7, fig.height=5, eval = F}
x <- df3[dataset_calibration_scale == "WMO CO2 X2019"]
x$co2 <- x$value*1e+6
obs_plot(x, 
         time = "co2", 
         y = "altitude_final", 
         colu = "site_code", #n = c(1L, 3L, 6L, 8L, 9L, 11L, 12L), 
         type = "p", 
         xlab = expression(CO[2]~ppm), 
         ylab = "altitude (m)")
```
